The \emph{Photochem} photochemical model

### 2023-05-18 ms.tex
tinyu

\begin{align}
  f_i^{j} &\approx \overline{f_i}^{j} \nonumber \\
  n^{j} &\approx \overline{n}^{j} \nonumber \\
  \sigma_i^j &\approx \overline{\sigma_i}^j(t) \nonumber \\
  \Phi_{i}^{j+1/2} &\approx \Phi_{i}(z^{j+1/2},t) \nonumber \\
  \Phi_{i}^{j-1/2} &\approx \Phi_{i}(z^{j-1/2},t) \nonumber
\end{align}
Re-writing Equation \eqref{eq:continuity_simple_conservative1}, substituting in the approximate quantities gives

@@

In the equation for gas flux (Equation \eqref{eq:phi_gas_1}), the first term is diffusive and the second term is advective. For most all plausible diffusion coefficients and advection velocities the diffusion term dominates, therefore, it is suitable to use a centered finite volume scheme for $\Phi_{i\text{,gas}}^{j+1/2}$ and $\Phi_{i\text{,gas}}^{j-1/2}$ (Equation 4.10 in \cite{Leveque_2002}):

@@

\begin{equation}
  \frac{\partial f_{i\in\text{particle}}^j}{\partial t} = - \frac{1}{n^{j}} \frac{\Phi_{i\text{,diff}}^{j+1/2} - \Phi_{i\text{,diff}}^{j-1/2}}{\Delta z^j} - \frac{1}{n^{j}} \frac{\Phi_{i\text{,advec}}^{j+1/2} - \Phi_{i\text{,advec}}^{j-1/2}}{\Delta z^j} + \frac{\sigma_i^j}{n^{j}}
\end{equation}
Applying a centered scheme to the diffusive fluxes follows the same procedure as we did with gas diffusion (Section \ref{sec:FV_particles}). To apply a first-order upwind scheme, we first recognize that falling particles means information is traveling from up to down in the model. Therefore, an upwind scheme approximates advective fluxes with a forward difference

@@

Equations \eqref{eq:dfdt_general} and \eqref{eq:dfdt_boundary} are a system of ordinary differential equations (ODEs) which are a finite volume approximation of our original system of PDEs (Equation \eqref{eq:continuity_simple}). The number of ODEs is the number evolving species multiplied by the number of atmospheric grid-cells. We often use about 100 species and 200 vertical layers resulting in $\sim$20,000 ODEs.

@@

\begin{table}
  \centering
  \begin{tabularx}{\linewidth}{p{0.15\linewidth} | p{0.55\linewidth} | p{0.3\linewidth}}
  \hline \hline
  Variable & Definition & Units \\
  \hline
  \(f_{i}\) & Mixing ratio of species \(i\) & dimensionless \\
  \(n_{i}\) & Number density of species \(i\) & molecules
  cm\textsuperscript{-3} \\
  \(z\) & Altitude & cm \\
  \(t\) & Time & seconds \\
  \(n\) & Total number density & molecules
  cm\textsuperscript{-3} \\
  \(P_{i}\) & Total chemical production of species \(i\) & molecules
  cm\textsuperscript{-3} s\textsuperscript{-1} \\
  \(L_{i}\) & Total chemical loss of species \(i\) & molecules
  cm\textsuperscript{-3} s\textsuperscript{-1} \\
  \(R_{\text{i, rainout}}\) & Production and loss of species \(i\) from
  rainout & molecules cm\textsuperscript{-3}
  s\textsuperscript{-1} \\
  \(Q_{\text{i, lightning}}\) & Production and loss of species \(i\)
  from lightning & molecules cm\textsuperscript{-3}
  s\textsuperscript{-1} \\
  \(\Phi_{i}\) & Vertical flux of species \(i\) & molecules
  cm\textsuperscript{-2} s\textsuperscript{-1} \\
  \(K\) & Eddy diffusion coefficient & cm\textsuperscript{2}
  s\textsuperscript{-1} \\
  \(D_{i}\) & Molecular diffusion coefficient & cm\textsuperscript{2}
  s\textsuperscript{-1} \\
  \(H_{i}\) & \(= N_{a}\text{kT}\text{/}\mu_{i}g\), The scale heights of
  species \(i\) & cm \\
  \(H_{a}\) & \(= N_{a}\text{kT}\text{/}\overline{\mu}g\), The average
  scale height. & cm \\
  \(N_{a}\) & Avogadro's number & molecules
  mol\textsuperscript{-1} \\
  \(k\) & Boltzmann's constant & erg K\textsuperscript{-1} \\
  \(\mu\) & Molar mass. \(\overline{\mu}\) is mean molar mass of the
  atmosphere, and \(\mu_{i}\) is the molar mass of species \(i\) & g
  mol\textsuperscript{-1} \\
  \(g\) & Gravitational acceleration & cm
  s\textsuperscript{-2} \\
  \(\alpha_{\text{Ti}}\) & Thermal diffusion coefficient of species \(i\).
  We neglect this term (\(\alpha_{\text{Ti}} = 0\)) &
  dimensionless \\
  \(T\) & Temperature & K \\
  $w_i$ & Fall velocity of a particle & cm s$^{-1}$ \\
  $\eta$ & Dynamic viscosity of air. Equation \eqref{eq:dynamic_viscosity}. & dynes s cm$^{-2}$ \\
  \end{tabularx}
  \caption{Variables in \eqref{eq:continuity}}
  \label{tab:variables}
  \end{table}

>>>>

\begin{equation*}
  \begin{aligned}
    f_i^{j} &\approx \overline{f_i}^{j} \\
    n^{j} &\approx \overline{n}^{j} \\
    \sigma_i^j &\approx \overline{\sigma_i}^j(t)
  \end{aligned}
  \quad\quad\quad
  \begin{aligned}
    \Phi_{i}^{j+1/2} &\approx \Phi_{i}(z^{j+1/2},t) \\
    \Phi_{i}^{j-1/2} &\approx \Phi_{i}(z^{j-1/2},t)
  \end{aligned}
\end{equation*}
Re-writing Equation \eqref{eq:continuity_simple_conservative1}, substituting in the approximate quantities gives

@@

In the equation for gas flux (Equation \eqref{eq:phi_gas_1}), the first term is diffusive and the second term is advective. For most all plausible diffusion coefficients and advection velocities the diffusion term dominates, therefore, it is suitable to use a centered finite volume scheme (Equation 4.10 in \cite{Leveque_2002}):

@@

\begin{equation}
  \frac{\partial f_{i\in\text{particle}}^j}{\partial t} = - \frac{1}{n^{j}} \frac{\Phi_{i\text{,diff}}^{j+1/2} - \Phi_{i\text{,diff}}^{j-1/2}}{\Delta z^j} - \frac{1}{n^{j}} \frac{\Phi_{i\text{,advec}}^{j+1/2} - \Phi_{i\text{,advec}}^{j-1/2}}{\Delta z^j} + \frac{\sigma_i^j}{n^{j}}
\end{equation}
Applying a centered scheme to the diffusive fluxes follows the procedure shown in Section \ref{sec:FV_gases}. To apply a first-order upwind scheme, we first recognize that if particle are falling, then information is traveling from up to down in the model. Therefore, an upwind scheme approximates advective fluxes with a forward difference

@@

Equations \eqref{eq:dfdt_general} and \eqref{eq:dfdt_boundary} are a system of ordinary differential equations (ODEs) which are a finite volume approximation of our original system of PDEs (Equation \eqref{eq:continuity_simple}). The number of ODEs is the number evolving species (gases + particles) multiplied by the number of atmospheric grid-cells. We often use about 100 species and 200 vertical layers resulting in $\sim$20,000 ODEs.

@@

\begin{table}
  \centering
  \begin{tabularx}{\linewidth}{p{0.15\linewidth} | p{0.55\linewidth} | p{0.3\linewidth}}
  \hline \hline
  Variable & Definition & Units \\
  \hline
  \(f_{i}\) & Mixing ratio of species \(i\) & dimensionless \\
  \(n_{i}\) & Number density of species \(i\) & molecules
  cm\textsuperscript{-3} \\
  \(z\) & Altitude & cm \\
  \(t\) & Time & seconds \\
  \(n\) & Total number density & molecules
  cm\textsuperscript{-3} \\
  \(P_{i}\) & Total chemical production of species \(i\) & molecules
  cm\textsuperscript{-3} s\textsuperscript{-1} \\
  \(L_{i}\) & Total chemical loss of species \(i\) & molecules
  cm\textsuperscript{-3} s\textsuperscript{-1} \\
  \(R_{\text{i, rainout}}\) & Production and loss of species \(i\) from
  rainout & molecules cm\textsuperscript{-3}
  s\textsuperscript{-1} \\
  \(Q_{\text{i, lightning}}\) & Production and loss of species \(i\)
  from lightning & molecules cm\textsuperscript{-3}
  s\textsuperscript{-1} \\
  \(\Phi_{i}\) & Vertical flux of species \(i\) & molecules
  cm\textsuperscript{-2} s\textsuperscript{-1} \\
  \(K\) & Eddy diffusion coefficient & cm\textsuperscript{2}
  s\textsuperscript{-1} \\
  \(D_{i}\) & Molecular diffusion coefficient & cm\textsuperscript{2}
  s\textsuperscript{-1} \\
  \(H_{i}\) & \(= N_{a}\text{kT}\text{/}\mu_{i}g\), The scale heights of
  species \(i\) & cm \\
  \(H_{a}\) & \(= N_{a}\text{kT}\text{/}\overline{\mu}g\), The average
  scale height. & cm \\
  \(N_{a}\) & Avogadro's number & molecules
  mol\textsuperscript{-1} \\
  \(k\) & Boltzmann's constant & erg K\textsuperscript{-1} \\
  \(\mu\) & Molar mass. \(\overline{\mu}\) is mean molar mass of the
  atmosphere, and \(\mu_{i}\) is the molar mass of species \(i\) & g
  mol\textsuperscript{-1} \\
  \(g\) & Gravitational acceleration & cm
  s\textsuperscript{-2} \\
  \(\alpha_{\text{Ti}}\) & Thermal diffusion coefficient of species \(i\).
  We neglect this term (\(\alpha_{\text{Ti}} = 0\)) &
  dimensionless \\
  \(T\) & Temperature & K \\
  $w_i$ & Fall velocity of a particle & cm s$^{-1}$ \\
  $\eta$ & Dynamic viscosity of air. Equation \eqref{eq:dynamic_viscosity}. & dynes s cm$^{-2}$ \\
  \end{tabularx}
  \caption{Variables used in the text}
  \label{tab:variables}
  \end{table}
